Evaluation of dynamic resource allocation strategies for overlay Apache Spark clusters on HPC

### 2019-05-01 paper.tex
some fixes to intro

\begin{document}
\title{SPA: An Apache Spark Pilot scheduler\\ for HPC applications}
\author{
    \IEEEauthorblockN{
        Val\'erie Hayot-Sasson and Tristan Glatard
    }
    \IEEEauthorblockA{}
}
\maketitle

@@

In the recent years, neuroimaging \TG{If you want to put neuroimaging first, then you need to explain it. Just say that it's an example among many fields}
Big Data has become more accessible, as a
result of the growing number of open-data initiatives~\cite{openneuro, hcp, ukbiobank}. However, the
processing of such large datasets using standard neuroimaging pipelines has 
become cumbersome. As a result, many researchers are moving towards Big Data 
engines, such as Apache Spark~\cite{spark} and Dask~\cite{sdask}, to 
create their processing pipelines. Unlike traditional neuroimaging 
engines, Big Data engines were designed to be executed on dedicated 
commodity infrastructures using Big Data schedulers. However, 
researchers still largely rely on High Performance Computing (HPC) 
infrastructure to process their data, although they may have access to 
clouds or local workstations.

High-Performance Computing infrastructure differs from the commodity infrastructure 
typically used by Big Data platforms in that it is not dedicated. That is, there 
are multiple users running a variety of different applications and all using the 
same resources. Schedulers used on HPC infrastructure include PBS, SGE, Slurm, HTCondor and 
TORQUE~\cite{schedulers}. The majority of these are batch submission schedulers
in which the number of resources are requested by the user and the specified 
program is executed on the infrastructure.

Another popular HPC scheduling approach,
implemented notably by HTCondor~\cite{htcondor}, is known as pilot scheduling. Rather than requesting
all necessary resources in a single batch call, a pilot-scheduling system will request
multiple instances of subsets of the required resources. Processing starts 
as soon as a minimal amount of resources are allocated, and can be replenished if 
resources are lost. Such a scheduling approach decreases the time a 
user spends in the queue waiting for resources as less resources are 
being requested per instance.\todo{need citation} In addition, pilot 
scheduling makes it easier for users to estimate the number of 
cores required by their application, as a slight overestimation would not impact queuing 
time. Many scientific workflow engines are compatible with 
pilot-scheduling schedulers~\cite{nipype and others} and some others 
exclusively use pilot-scheduling approaches~\cite{Pegasus and PSOM}. 
\tristan{I'm not sure if Pegasus is exclusively pilot based.}

Apache Spark is a popular Big Data framework, commonly used in both industrial
and academic settings. Although it is a Scala-based framework, it also has APIs
for Java, Python (PySpark) and R. Spark's Resilient Distributed Dataset (RDD) abstraction enabled
in-memory processing of pipelines by co-locating tasks and data, which provided important performance improvements compared to its predecessors.
 Through the use of RDDs, it also became possible to execute iterative 
workflows -- something not easily doable in older frameworks such as Hadoop MapReduce.
Schedulers for Spark include its built-in standalone schedule, Yet Another Resource Negotiator (YARN),
and Mesos~\cite{yarn, mesos}. Out of the three schedulers, only Mesos can be used
as an HPC scheduler, which is not frequently done. \tristan{why?}

@@

There have been some efforts to use pilot-scheduling approaches with Spark applications
running on HPC \cite{jha and spark on pbs paper}. However, these projects are either
not open-source or not necessarily compatible with Slurm.

>>>>

\begin{document}
\title{Evaluation of dynamic resource allocation strategies for overlay Apache
       Spark clusters on HPC}
\author{
    \IEEEauthorblockN{
        Val\'erie Hayot-Sasson and Tristan Glatard
    }
    \IEEEauthorblockA{}
}
\maketitle

@@

With the increasing volume of data being collected and shared, Big Data has become
an important area of focus to many scientific domains. As is the case with neuroimaging,
dataset available publicly now reach up to Petabytes in size~\cite{UKBiobank, hcp}. Average workstations found in research lab would not be
able to support the processing of such large dataset, leaving researchers to look 
into cloud or High-Performance Computing~(HPC) solutions. Whereas both options are
suitable for the processing of scientific Big Data, HPC clusters remain the more 
sought after solution, potentially due to its cost-effectiveness.

Scientific workflows may be quite compute intensive, requiring more than a few
hours to complete (\cite{freesurfer, fmriprep, a paper which mentions duration}).
The coupling of large datasets with compute-intensive workflows may result in 
unreasonably long processing times. Such limitations restrict the amount data 
utilized in scientific experiments. To reduce the effects of Big Data on processing
times, researchers have begun to move towards employing Big Data frameworks for their
analyses. Examples of Big Data frameworks including Hadoop MapReduce~\cite{Hadoop},
Apache Spark~\cite{Spark} and Dask~\cite{Dask}. However, these frameworks were 
designed with dedicated commodity infrastructure in mind, and, with the exception
of Dask, do not support traditional HPC schedulers.

Within High-Performance Computing clusters, many users processing a variety of different
applications, share resources. These shared resources include storage (e.g. Parallel file system) and 
compute node resources (memory, cpu, local disk space, etc.). In order for users to 
gain access to shared compute resources, HPC clusters typically employ schedulers for 
batch resource provisioning. Examples of such schedulers include PBS, SGE, Slurm and 
TORQUE~\cite{schedulers}. These schedulers require that the user specify the amount
of resources required and the duration that they require the resources for. The users 
are then placed in a queue at a position based on the user's priority and resource 
availability.

There are several limitations to running applications on batch HPC schedulers, 
particularly in the face of Big Data. For instance, users may not necessarily 
know the exact requirements of the applications. With applications that are 
both data- and compute-intensive, this may result in the user overestimating the
resources required and the duration that these resources are required for. Frequent
high overestimates may create bottlenecks in the system, preventing other users
from occupying allocated idle resources. In contrast, users may underestimate required 
resources. This may subsequently result in many failures within the application leading to
additional re-executions of the same application. This is problematic for the user
as their priority may be affected by the cluster usage. Additionally, the user may 
require more time on the cluster to relaunch non-checkpointed and interrupted 
tasks. Furthermore, Big Data frameworks, such as Apache Spark, do not interact naturally
with HPC schedulers and typically rely on other schedulers for resource provisioning
and task scheduling. Therefore, to run Big Data applications on HPC schedulers, 
it is also necessary to start an overlay cluster.

In contrast to static resource provisioning, as is the case with batch HPC schedulers, 
dynamic resource provisioning, known as pilot job or glide-in scheduling, exists. 
This dynamic resource provisioning allows the user to specify a minimum and maximum 
amount of resources, and the cluster can expand and contract depending on application requirements.
Each pilot may consist of the minimal amount of resources required to launch the 
application. As each pilot would request allocation of fewer resources than that of a batch
request, Pilot Jobs would typically spend less time on the resource allocation 
queue. Furthermore, although individual pilots may expire due to walltimes, the dynamic
nature of pilots allows the entire application to surpass walltime limitations, as
additional pilots may be added to the pilot queue until application completion. 
Examples of HPC schedulers employing pilot strategies include HTCondor~\cite{htcondor}
and DIRAC~\cite{DIRAC}. \todo{maybe mention PSOM and Pegasus here}

Apache Spark is a popular Big Data framework, commonly used in both industrial
and academic settings. Although it is a Scala-based framework, it also has APIs
for Java, Python (PySpark) and R. Spark's Resilient Distributed Dataset (RDD) abstraction enabled
in-memory processing of pipelines by co-locating tasks and data, which provided important performance improvements compared to its predecessors.
 Through the use of RDDs, it also became possible to execute iterative 
workflows -- something not easily doable in older frameworks such as Hadoop MapReduce.
Schedulers for Spark include its built-in standalone schedule, Yet Another Resource Negotiator (YARN),
and Mesos~\cite{yarn, mesos}.

@@

Pilot scheduling has been applied to scientific workflow engines including Pegasus,
PSOM and RADICAL-Pilot. While there have been some efforts on implementing pilot scheduling
approaches for \cite{jha and spark on pbs paper}. However, these projects are either
not open-source or not necessarily compatible with Slurm.
